ProbabilityBoundsAnalysis.jl:\\ Arithmetic with sets of distributions

### 2022-04-14 paper/paper.tex
shortens paper

% JuliaCon proceedings template
\documentclass{juliacon}
\setcounter{page}{1}
\usepackage{caption}
\usepackage{amsmath}
\usepackage{xcolor}
%\bibliographystyle{juliacon}
\usepackage[labelformat=simple]{subfig}
\renewcommand\thesubfigure{\arabic{subfigure}}
\renewcommand{\subfigurename}{Figure}
\begin{document}
\input{header}
\maketitle
\begin{abstract}
Probability bounds analysis combines interval arithmetic with probability theory, and provides a representation of sets of distributions in structures called probability boxes (p-boxes). P-boxes generalise both distribution functions and intervals, and return interval bounds on all probabilistic quantities, for example sample realisations, cdfs, and probability measures are all intervals. This framework also allows for the comprehensive propagation of probabilities through calculations in a rigorous way, in a similar fashion that interval arithmetic does for sets of real values. \texttt{ProbabilityBoundsAnalysis.jl} provides a rigorous arithmetic of random variables, where both marginal (univariate) distributions and dependency information can be known, partially known or missing entirely. We describe the main theoretical elements of probability bounds analysis, and provide a simplified implementation of the method in code snippets which can be readily evaluated in the Julia command terminal.
\end{abstract}
\section{Introduction}
\label{sec:intro}
An arithmetic of probability distributions has long held interest among mathematicians and scientists. Indeed it was Kolmogorov who originally asked about the result of an operation between two random variables without knowing their joint distribution. This was answered for the sum by Makarov \cite{makarov1982estimates}, who showed the result was a set of distributions and was able to provide bounds on this function. Sklar, Schweizer and Frank generalised this result to other positive binary operations \cite{frank1987best,schweizer2011probabilistic}. In this pursuit they created copulas, a general way to encode probabilistic dependence independently from marginals. These are now an essential object used in probabilistic modelling. In his dissertation \cite{williamson1989probabilistic}, Williamson described an algorithm for efficiently performing these arithmetic operations, which can give guaranteed bounds on probability distributions in terms of an upper and lower cumulative distribution function (cdf). He called his method \textit{probabilistic arithmetic} and his sets of distributions \textit{dependency bounds}. Since then the method has been generalised \cite{ferson2015constructing,ferson1996whereof,ferson2004arithmetic} to most of the base binary and unary operations that would be present in a programming language. Probability boxes (p-boxes) are the name now given to these structures.

@@

The idea of bounding probability has a long tradition throughout the history of probability theory. George Boole \cite{boole1854investigation, hailperin1986boole} used the notion of interval bounds on probability. Chebyshev \cite{chebyshev1874valeurs} described bounds on a distribution when only the mean and variance of the variable are known, and Markov \cite{markoff1900question} found bounds on a positive variable when only the mean is known.  Fréchet \cite{frechet1935generalisation} discovered how to bound joint distributions solely from knowing the marginal distributions, without making independence assumptions. Bounding probabilities has continued to the present day, culminating into the modern theory of imprecise probabilities \cite{walley1991statistical, klir2013uncertainty, troffaes2014lower, augustin2014introduction}.

Imprecise probability is effectively a generalisation of probability theory where uncertainty can be expressed about the probability measure. This is particularly relevant when information is scarce, unreliable, vague, conflicting or imprecise. In such cases defining a unique probability distribution is difficult. P-boxes are one of many ways to describe a set of distributions. Others include: Dempster-Shafer structures \cite{dempster2008upper,shafer1976mathematical}, random sets \cite{molchanov2005theory}, possibility distributions \cite{zadeh1978fuzzy,dubois1988possibility, hose2019possibilistic}, lower previsions \cite{troffaes2014lower}, and credal sets \cite{levi1983enterprise}. These structures were discovered independently, but are often synonymous and can be translated from one to another, with different degrees of generality. Imprecise probability unifies the underlying theories. For a comprehensive overview of the theory, and a formal description of uncertainty and information in terms of these structures, \cite{klir2013uncertainty} is recommended.

@@

\end{figure*}
\subsection{Automatic uncertainty in a programming language: why Julia?}
The long term goal of probability bounds analysis is to create a fully uncertain programming language, where any computer variable may be represented as an interval, distribution, p-box or other imprecise quantity. Such a framework would allow for uncertain extensions of deterministic functions to be computed in an automatic, rigorous, and tight fashion. Those with little knowledge of probability theory could add uncertainty to their computations with minimal effort. Stochastic computation often require specialist training to be used effectively, and are usually not rigorous and are famously computationally expensive. \texttt{ProbabilityBoundsAnalysis.jl} provides the arithmetic basis for such a framework. In this section we give some arguments why Julia \cite{bezanson2017julia} is an ideal target language for such a framework.

@@

\subsubsection{Metaprogramming}
An uncertain programming language would require \textit{source-to-source} transformations at times. For example, one way to evaluate a programs control-flow (\texttt{if-else} conditions) with uncertain inputs is to use metaprogramming. The programs control flow could be unraveled, and with each possible conditional branch evaluated for the possible instance of the variable, and then combined as an output. For example consider the following function
\begin{lstlisting}[language = Julia]
  function unc_flow(x)
    if x < 0 
      return x - 1
    else 
      return x
    end
  end
\end{lstlisting}
\noindent An interval evaluation of $\text{unc\_flow}([-1,1])$ would be the union of the output of the different branches, the interval union $[-2, -1) \cup [0, 1]$. For a language like Python, this would require an uncertainty pre-compiler (like that proposed in \cite{gray2019computing}) whereby the python code is first parsed using a separate program, and then evaluated in Python. In Julia, this can be handled natively with its metaprogramming capabilities.
\begin{figure*}[htp]

@@

\fi
\section{Probability Boxes}
\label{sec:pboxes}
A real-valued random variable is characterised by its distribution function $F$, which is a monotonically increasing function from the real numbers onto the interval $[0, 1]$ such that the value of the function at negative infinity is zero and the value of the function at positive infinity is one.  A p-box consists of a pair of such functions that are used to circumscribe an imprecisely known distribution function $F$. The p-box, in its simplest form consisting of the pair of bounding functions, identifies a set of probability distributions that lie entirely within these bounds. Additional information about the random variable may be available, such as bounds on its mean and variance and its distribution family, which may be used to further restrict the set of distributions. A p-box is thus defined by the following three constraints: (1) two bounding cumulative distribution functions (cdf), (2) interval bounds on the mean and variance, and (3) a collection of distribution families:
\begin{enumerate}
  \item $\begin{aligned}[t]
    \underline{F}(x) \leq F(x) \leq \overline{F}(x), \\%&&&& \text{(Bounds on the cdf)} \\
  \end{aligned}$
  \item $\begin{aligned}[t]
    \mu &\in [\underline{ \mu }, \overline{ \mu }]  \\%&&&& \text{(bounds on mean and variance)}\\
    \sigma^2 &\in [\underline{\sigma}^2 , \overline{\sigma}^2],
  \end{aligned}$
  \item $\begin{aligned}[t]
      F \in \bm{F}. \\%&&&&&&& \text{(distribution family)} \\
  \end{aligned}$
  \end{enumerate}
\noindent That is, a random variable is a member of a p-box if its cdf $F$ falls within the cdf bounds of the p-box $F(x) \in [\underline{F}(x), \overline{F}(x)]$ for all $x$, its moments are inside the interval moments of the p-box, and it belongs to a family of distribution functions (e.g. normal, uniform) considered by the p-box. Some of the constraints may be missing. For example, if the distribution family is unknown then the set is defined solely from the cdf and moment bounds. Such p-boxes are sometimes called non-parametric, since its members do not belong to any particular class of distribution. Some constraints may also be inferred from others. For example, the interval moments may be bounded from the cdf bounds, and cdf may be bounded from moment information (explored further in section \ref{section:Moments}). Figure \ref{fig:figure1} (scripts for figures found in \texttt{/examples/JuliaCon}) shows an example of a distribution function, interval and a p-box. The grey shaded region in the interval and p-box bounds an infinite collection of probability distributions.

@@

  \caption{Various parametrically constructed p-boxes with interval parameters.}
  \label{fig:figure3}
\end{figure*}
P-boxes generalise precise distributions and intervals in the following way. A distribution is a p-box with a precise cdf and moments, i.e. when
\begin{align*}
  \underline{F}(x) &= \overline{F}(x), \\ 
  \underline{\mu}  &= \overline{\mu}, \\ 
  \underline{\sigma}^2 &= \overline{\sigma}^2 .
\end{align*}
\noindent An interval $[a,b]$ is a p-box whose bounds are step functions
\begin{align*}
    \underline{F}(x) &= \epsilon_{b}(x) ,\\
    \overline{F}(x) &= \epsilon_{a}(x) ,
\end{align*}
\noindent where $\epsilon_k$ is
\begin{equation*}
   \epsilon_k(x) = \begin{cases} 0 &\text{when } x < k \\ 1 &\text{when } x \geq k. \end{cases} 
\end{equation*}
\noindent Moreover, theoretical bounds on the mean and variance of an interval can be found \cite{ferson2002ramas}
\begin{align*}
  \mu &\in [a, b], \\
  \sigma^2 &\in [0, (b - a)^{2}/4].
\end{align*}
\noindent That is, it is not possible to find a distribution inside $[a, b]$ with a variance greater than $(b-a)^{2}/4$. The lower bound on the variance is zero, since scalars are also included in the interval. Under this definition of an interval, a random sample will always be the interval $[a, b]$, the cdf returns 0 when $x < a$, the interval $[0,1]$ when $a \leq x < b$, and $1$ when $b \leq x$. Further, the probability measure of the interval $X = [a, b]$ will be: 
\begin{equation*}
  \mathbb{P}_{X}(U) = \begin{cases}
    0 & \text{when } U \cap [a,b] = \emptyset \\
    1 & \text{when } [a,b] \subseteq U  \\
    [0, 1] & \text{otherwise.}\\
  \end{cases}
\end{equation*}
\noindent That is, if the set $U$ does not intersect the interval $X$, $\mathbb{P}_{X}=0$, i.e. $U$ certainly does not contain any of the random variables. If $X$ is fully contained in $U$, $\mathbb{P}_{X}=1$, i.e. $U$ certainly contains all of the random variables. And finally if $U$ intersects (but does not contain) $X$, the probability measure is the vacuous probability interval $\mathbb{P}_{X} = [0, 1]$, i.e. we are completely uncertain about the containment. Note that the $P_{X} = \{0, 1, [0,1]\}$ is due to the interval bounds being step functions. Generally p-boxes may yield other probability intervals.
\subsection{Outer representations of p-boxes}
\label{sec:outer_approx}
An important feature of probability bounds analysis is how distribution functions and p-boxes are represented. Generally, analytical solutions for cdfs are not readily available, for example the normal distribution's cdf can only be found by integrating the probability density, usually with quadrature. However, even if the functions were available analytically, when the variables are used in an arithmetic operation the output distribution will not necessarily belong to the same family. Therefore, we require a representation of these continuous functions which is not dependent on the distribution family (for example as used in polynomial chaos expansion), and ideally is robust in the sense that it can produce an interval error for the numerical representation. Such a representation of a probability is sometimes called an "outer approximation", in the same fashion an interval is an outer approximation of a floating point number. Such outer approximations contract to the exact value or function as more computational resources are used.

We follow the representation first introduced by Williamson and Downs \cite{williamson1990probabilistic}, where a discrete upper and lower approximations of distribution functions are constructed using the inverse cdfs. Note that when considering inverse cdfs $\underline{F}^{-1}(p) \geq F^{-1}(p) \geq \overline{F}^{-1}(p)$, for $p \in [0,1]$. Beginning vector indexing by $1$, an outer approximation using two finite vectors $u$ and $d$ of length $N$ is constructed by evaluating the inverse cdfs for uniformly spaced probability levels $p_{i} = \frac{i-1}{N}$ for $i = 1, ... , N+1$, e.g. $p_{i} = [0, 0.25, 0.5, 0.75, 1]$ for $N = 4$. The vectors $u$ and $d$ are defined as follows
\begin{align*}
  u[i] &= \overline{F}^{-1}(p_{i}), \\ 
  d[i] &= \underline{F}^{-1}(p_{i+1}), & \text{for } i = 1, ..., N .
\end{align*}
The vectors $u$ and $d$ are a collection of "real" or "physical" evaluations of a the random variable $X$, as opposed to cdf evaluations. Although the probability levels $p$ are uniformly spaced, $u$ and $d$ are generally not. The vectors $u$ and $d$ can then serve as bounds on the monotonic functions $\overline{F}^{-1}$ and $\underline{F}^{-1}$
\begin{align*}
  u[i] &\leq \overline{F}^{-1}(p), & p_{i} \in [p_{i}, p_{i+1} ),\\ 
  d[i] &\geq \underline{F}^{-1}(p), & p_{i} \in (p_{i}, p_{i+1}].
\end{align*}
That is, in the region $[p_{i}, p_{i+1})$, $u[i]$ is a lower bound on $\overline{F}^{-1}$ and $d[i]$ is an upper bound on $\underline{F}^{-1}$ in $(p_{i}, p_{i+1}]$. In the case that the p-box characterises a precise distribution and $\overline{F}^{-1}(p) = \underline{F}^{-1}(p)$ gives:  $u[i+1] = d[i]$. Figure \ref{fig:figure2} shows examples of a distribution and a p-box with various outer representations.

@@

  struct pbox{T <: Real}
    u :: Vector{T}    # left discrete inverse
    d :: Vector{T}    # right discrete inverse
    m :: Interval{T}  # interval mean
    v :: Interval{T}  # interval variance
    shape :: String   # shape
  end
\end{lstlisting}
\noindent where the vectors $u$ and $d$ are the "up" and "down", or the "left" and "right", edges of the p-box. There can be confusion since $\overline{F}$ is the upper bound on the cdf, but $\overline{F}^{-1}$ is the lower bound on the random variable, i.e. the left edge of the p-box. We therefore call $\overline{F}$ the "up" or "left" edge, and $\underline{F}$ the "down" or the "right" edge. The intervals $m$ and $v$ are the interval moments, and the $shape$ is a string containing any distribution family information. An example of a normal distribution constructor is as follows

\begin{lstlisting}[language = Julia]
  using Distributions
  const steps = 200

@@

    return interval(lb, ub)
  end
\end{lstlisting}
\noindent Other available functions on p-boxes which we do not describe here are
\begin{itemize}
  \item \textbf{env} or $\cup$: takes the envelope or hull of two p-boxes. A convex set-union of the probability sets,
  \item \textbf{imp} or $\cap$: imposition or intersection of two p-boxes. Set-intersection of probability sets,
  \item \textbf{subseteq} or $\subseteq$: checks if a p-box is a subset or equal to another,
  \item \textbf{in} or $\in$: checks if a distribution is a member of a p-box,
  \item \textbf{mixture}: a stochastic mixture of p-boxes.
\end{itemize}
\subsection{Where do p-boxes come from?}
\noindent In this section we discuss some situations where p-boxes arise naturally.
\subsubsection{Partial distributional information}
In situations where the parametric family of the uncertainty is known, for example it is known that $X$ is normally distributed $X \sim N(\mu_{X}, \sigma_{X}^{2})$ or that $Y$ is a uniform $Y \sim U(a_{Y},b_{Y})$, but the precise parameters to these functions are unknown, then p-boxes are a natural representation. In the case that bounds are known on parameters, then bounds on the cdf can be found using the interval parameters and the inverse cdf of the parametric family. Figure \ref{fig:figure3} shows a variety of p-boxes constructed this way.
\subsubsection{Missing family but partial moment information}
\label{section:Moments}
It is well known that generally a random variables distribution function cannot be uniquely determined solely from its moment information, even with an infinite number of moments, unless some assumptions are made about the family (i.e. it's normal). In probability theory this is known as the \textit{moment problem}. However, in these situations it is possible to define a p-box which bounds all random variables with the known properties, without loss of generality or making unwarranted assumptions.

These p-boxes are constructed using the Chebyshev (mean and variance) \cite{chebyshev1874valeurs}, Markov (mean and one bound) \cite{markoff1900question} and Cantelli (mean and range) inequalities from probability theory. These inequalities allow for the cdf to be bounded from various range and moment information. For example, the Chebyshev inequality can be used to find p-box bounds from the random variables mean and variance
\begin{equation*}
  \underline{F}_{X}(x) = \begin{cases} 1 / (1 + (x - \mu_{X})^2 / \sigma^{2}_{X}),& \text{if } \mu_{X}<x \\
    1,& \text{if } x\leq \mu_{X} \end{cases}
\end{equation*}
\begin{equation*}
  \overline{F}_{X}(x)  = \begin{cases} 1 / (1 + \sigma^{2}_{X}/(x - \mu_{X})^2),& \text{if } x<\mu_{X} \\
  1,& \text{if } \mu_{X}\leq x \end{cases}
\end{equation*}
\noindent A list of available distribution-free constructors are
\begin{itemize}
  \item \textbf{MeanVar (or MeanStd)}: using Chebyshev,
  \item \textbf{MeanMin (or MeanMax)}: using Markov,
  \item \textbf{MeanMinMax}: using Cantelli,
  \item \textbf{MinMaxMeanVar (or MinMaxMeanStd)}: using \cite{ferson2021distribution}
\end{itemize}
\noindent All distribution-free constructors may also take interval moments parameters.
\subsubsection{Operations involving intervals and distributions}
An arithmetic operation involving an interval and a precise distribution will produce set of distributions, i.e. a unique distribution for every real number in the interval, whose p-box bounds can be found. The following function performs a binary operation between a p-box and an interval
\begin{lstlisting}[language = Julia]
  using MomentArithmetic

@@

A method which is growing in popularity are inferential models \cite{martin2015inferential}. These generalise classical inference methods to account for limited statistical knowledge, and give a statistical inference theory which is non-additive (as opposed to classical probability which is additive). These inference procedures, often based on random sets, give certain validity and performance guarantees. Inferential models can viewed from the perspective of bounding sets of distributions, and is closely linked to possibility theory \cite{liu2020inferential}. A p-box can be prescribed to such structures, leading to structures like confidence boxes or c-boxes \cite{fersoncomputing}, which follow the same arithmetic rules as p-boxes.

@@

%\section{Operations on p-boxes}

%\section{P boxes}

>>>>

% JuliaCon proceedings template
\documentclass{juliacon}
\setcounter{page}{1}
\usepackage{caption}
\usepackage{amsmath}
\usepackage{xcolor}
%\bibliographystyle{juliacon}
\usepackage[labelformat=simple]{subfig}
\renewcommand\thesubfigure{\arabic{subfigure}}
\renewcommand{\subfigurename}{Figure}
\begin{document}
\input{header}
\maketitle
\begin{abstract}
Probability bounds analysis combines interval arithmetic with probability theory, and provides a representation of sets of distributions in structures called probability boxes (p-boxes). P-boxes generalise both distribution functions and intervals, and return interval bounds on all probabilistic quantities, for example sample realisations, cdfs, and probability measures are all intervals. This framework also allows for the comprehensive propagation of probabilities through calculations in a rigorous way, in a similar fashion that interval arithmetic does for sets of real values. \texttt{ProbabilityBoundsAnalysis.jl} provides a rigorous arithmetic of random variables, where both marginal (univariate) distributions and dependency information can be known, partially known or missing entirely. We describe the main theoretical elements of probability bounds analysis, and provide a simplified implementation of the method in code snippets which can be readily evaluated in the Julia command terminal.
\end{abstract}
\section{Introduction}
\label{sec:intro}
An arithmetic of probability distributions has long held interest among mathematicians and scientists. Indeed it was Kolmogorov who originally asked about the result of an operation between two random variables without knowing their joint distribution. This was answered for the sum by Makarov~\cite{makarov1982estimates}, who showed the result was a set of distributions and was able to provide bounds on this function. Sklar, Schweizer and Frank generalised this result to other positive binary operations~\cite{frank1987best,schweizer2011probabilistic}. In this pursuit they created copulas, a general way to encode probabilistic dependence independently from marginals. These are now an essential object used in probabilistic modelling. In his dissertation~\cite{williamson1989probabilistic}, Williamson described an algorithm for efficiently performing these arithmetic operations, which can give guaranteed bounds on probability distributions in terms of an upper and lower cumulative distribution function (cdf). He called his method \textit{probabilistic arithmetic} and his sets of distributions \textit{dependency bounds}. Since then the method has been generalised~\cite{ferson2015constructing,ferson1996whereof,ferson2004arithmetic} to most of the base binary and unary operations that would be present in a programming language. Probability boxes (p-boxes) are the name now given to these structures.

@@

The idea of bounding probability has a long tradition throughout the history of probability theory. George Boole~\cite{boole1854investigation, hailperin1986boole} used the notion of interval bounds on probability. Chebyshev~\cite{chebyshev1874valeurs} described bounds on a distribution when only the mean and variance of the variable are known, and Markov~\cite{markoff1900question} found bounds on a positive variable when only the mean is known.  Fréchet~\cite{frechet1935generalisation} discovered how to bound joint distributions solely from knowing the marginal distributions, without making independence assumptions. Bounding probabilities has continued to the present day, culminating into the modern theory of imprecise probabilities~\cite{walley1991statistical, klir2013uncertainty, troffaes2014lower, augustin2014introduction}.

Imprecise probability is effectively a generalisation of probability theory where uncertainty can be expressed about the probability measure. This is particularly relevant when information is scarce, unreliable, vague, conflicting or imprecise. In such cases defining a unique probability distribution is difficult. P-boxes are one of many ways to describe a set of distributions. Others include: Dempster-Shafer structures~\cite{dempster2008upper,shafer1976mathematical}, random sets~\cite{molchanov2005theory}, possibility distributions~\cite{zadeh1978fuzzy,dubois1988possibility, hose2019possibilistic}, lower previsions~\cite{troffaes2014lower}, and credal sets~\cite{levi1983enterprise}. These structures were discovered independently, but are often synonymous and can be translated from one to another, with different degrees of generality. Imprecise probability unifies the underlying theories. For a comprehensive overview of the theory, and a formal description of uncertainty and information in terms of these structures, \cite{klir2013uncertainty} is recommended.

@@

\end{figure*}
\subsection{Automatic uncertainty in a programming language: why Julia?}
The long term goal of probability bounds analysis is to create a fully uncertain programming language, where any computer variable may be represented as an interval, distribution, p-box or other imprecise quantity. Such a framework would allow for uncertain extensions of deterministic functions to be computed in an automatic, rigorous, and tight fashion. Those with little knowledge of probability theory could add uncertainty to their computations with minimal effort. Stochastic computation often require specialist training to be used effectively, and are usually not rigorous and are famously computationally expensive. \texttt{ProbabilityBoundsAnalysis.jl} provides the arithmetic basis for such a framework. In this section we give some arguments why Julia~\cite{bezanson2017julia} is an ideal target language for such a framework.

@@

\subsubsection{Metaprogramming}
An uncertain programming language would require \textit{source-to-source} transformations at times. For example, one way to evaluate a programs control-flow (\texttt{if-else} conditions) with uncertain inputs is to use metaprogramming. The programs control flow could be unraveled, and with each possible conditional branch evaluated for the possible instance of the variable, and then combined as an output. For example consider the following function
\begin{lstlisting}[language = Julia]
  function unc_flow(x)
    if x < 0 
      return x - 1
    else 
      return x
    end
  end
\end{lstlisting}
\noindent An interval evaluation of $\text{unc\_flow}([-1,1])$ would be the union of the output of the different branches, the interval union $[-2, -1) \cup [0, 1]$. For a language like Python, this would require an uncertainty pre-compiler (like that proposed in~\cite{gray2019computing}) whereby the python code is first parsed using a separate program, and then evaluated in Python. In Julia, this can be handled natively with its metaprogramming capabilities.
\begin{figure*}[htp]

@@

\fi
\section{Probability Boxes}
\label{sec:pboxes}
A real-valued random variable is characterised by its distribution function $F$, which is a monotonically increasing function from the real numbers onto the interval $[0, 1]$ such that the value of the function at negative infinity is zero and the value of the function at positive infinity is one.  A p-box consists of a pair of such functions that are used to circumscribe an imprecisely known distribution function $F$. The p-box, in its simplest form consisting of the pair of bounding functions, identifies a set of probability distributions that lie entirely within these bounds. Additional information about the random variable may be available, such as bounds on its mean and variance and its distribution family, which may be used to further restrict the set of distributions. A p-box is thus defined by the following three constraints: (1) two bounding cumulative distribution functions (cdf), (2) interval bounds on the mean and variance, and (3) a collection of distribution families:
\begin{enumerate}
  \item $\begin{aligned}[t]
    \underline{F}(x) \leq F(x) \leq \overline{F}(x),
  \end{aligned}$
  \item $\begin{aligned}[t]
    \mu &\in [\underline{ \mu }, \overline{ \mu }]
    \sigma^2 &\in [\underline{\sigma}^2 , \overline{\sigma}^2],
  \end{aligned}$
  \item $\begin{aligned}[t]
      F \in \bm{F}.
  \end{aligned}$
\end{enumerate}
\noindent That is, a random variable is a member of a p-box if its cdf $F$ falls within the cdf bounds of the p-box $F(x) \in [\underline{F}(x), \overline{F}(x)]$ for all $x$, its moments are inside the interval moments of the p-box, and it belongs to a family of distribution functions (e.g. normal, uniform) considered by the p-box. Some of the constraints may be missing. For example, if the distribution family is unknown then the set is defined solely from the cdf and moment bounds. Such p-boxes are sometimes called non-parametric, since its members do not belong to any particular class of distribution. Some constraints may also be inferred from others. For example, the interval moments may be bounded from the cdf bounds, and cdf may be bounded from moment information (explored further in section \ref{section:Moments}). Figure \ref{fig:figure1} (scripts for figures found in \texttt{/examples/JuliaCon}) shows an example of a distribution function, interval and a p-box. The grey shaded region in the interval and p-box bounds an infinite collection of probability distributions.

@@

  \caption{Various parametrically constructed p-boxes with interval parameters.}
  \label{fig:figure3}
\end{figure*}
P-boxes generalise precise distributions and intervals in the following way. A distribution is a p-box with a precise cdf and moments, i.e. when
\begin{align*}
  \underline{F}(x) &= \overline{F}(x), \\ 
  \underline{\mu}  &= \overline{\mu}, \\ 
  \underline{\sigma}^2 &= \overline{\sigma}^2 .
\end{align*}
\noindent An interval $[a,b]$ is a p-box whose bounds are step functions
\begin{align*}
    \underline{F}(x) &= \epsilon_{b}(x) ,\\
    \overline{F}(x) &= \epsilon_{a}(x) ,
\end{align*}
\noindent where $\epsilon_k$ is
\begin{equation*}
   \epsilon_k(x) = \begin{cases} 0 &\text{when } x < k \\ 1 &\text{when } x \geq k. \end{cases} 
\end{equation*}
\noindent Moreover, theoretical bounds on the mean and variance of an interval can be found~\cite{ferson2002ramas}
\begin{align*}
  \mu &\in [a, b], \\
  \sigma^2 &\in [0, (b - a)^{2}/4].
\end{align*}
\noindent That is, it is not possible to find a distribution inside $[a, b]$ with a variance greater than $(b-a)^{2}/4$. The lower bound on the variance is zero, since scalars are also included in the interval. Under this definition of an interval, a random sample will always be the interval $[a, b]$, the cdf returns 0 when $x < a$, the interval $[0,1]$ when $a \leq x < b$, and $1$ when $b \leq x$. Further, the probability measure of the interval $X = [a, b]$ will be: 
\begin{equation*}
  \mathbb{P}_{X}(U) = \begin{cases}
    0 & \text{when } U \cap [a,b] = \emptyset \\
    1 & \text{when } [a,b] \subseteq U  \\
    [0, 1] & \text{otherwise.}\\
  \end{cases}
\end{equation*}
\noindent That is, if the set $U$ does not intersect the interval $X$, $\mathbb{P}_{X}=0$, i.e. $U$ certainly does not contain any of the random variables. If $X$ is fully contained in $U$, $\mathbb{P}_{X}=1$, i.e. $U$ certainly contains all of the random variables. And finally if $U$ intersects (but does not contain) $X$, the probability measure is the vacuous probability interval $\mathbb{P}_{X} = [0, 1]$, i.e. we are completely uncertain about the containment. Note that the $P_{X} = \{0, 1, [0,1]\}$ is due to the interval bounds being step functions. Generally p-boxes may yield other probability intervals.
\subsection{Outer representations of p-boxes}
\label{sec:outer_approx}
An important feature of probability bounds analysis is how distribution functions and p-boxes are represented. Generally, analytical solutions for cdfs are not readily available, for example the normal distribution's cdf can only be found by integrating the probability density, usually with quadrature. However, even if the functions were available analytically, when the variables are used in an arithmetic operation the output distribution will not necessarily belong to the same family. Therefore, we require a representation of these continuous functions which is not dependent on the distribution family (for example as used in polynomial chaos expansion), and ideally is robust in the sense that it can produce an interval error for the numerical representation. Such a representation of a probability is sometimes called an "outer approximation", in the same fashion an interval is an outer approximation of a floating point number. Such outer approximations contract to the exact value or function as more computational resources are used.

We follow the representation first introduced by Williamson and Downs~\cite{williamson1990probabilistic}, where a discrete upper and lower approximations of distribution functions are constructed using the inverse cdfs. Note that when considering inverse cdfs $\underline{F}^{-1}(p) \geq F^{-1}(p) \geq \overline{F}^{-1}(p)$, for $p \in [0,1]$. Beginning vector indexing by $1$, an outer approximation using two finite vectors $u$ and $d$ of length $N$ is constructed by evaluating the inverse cdfs for uniformly spaced probability levels $p_{i} = \frac{i-1}{N}$ for $i = 1, ... , N+1$, e.g. $p_{i} = [0, 0.25, 0.5, 0.75, 1]$ for $N = 4$. The vectors $u$ and $d$ are defined as follows
\begin{align*}
  u[i] &= \overline{F}^{-1}(p_{i}), \\ 
  d[i] &= \underline{F}^{-1}(p_{i+1}), & \text{for } i = 1, ..., N .
\end{align*}
The vectors $u$ and $d$ are a collection of "real" or "physical" evaluations of a the random variable $X$, as opposed to cdf evaluations. Although the probability levels $p$ are uniformly spaced, $u$ and $d$ are generally not. The vectors $u$ and $d$ can then serve as bounds on the monotonic functions $\overline{F}^{-1}$ and $\underline{F}^{-1}$
\begin{align*}
  u[i] &\leq \overline{F}^{-1}(p), & p_{i} \in [p_{i}, p_{i+1} ),\\ 
  d[i] &\geq \underline{F}^{-1}(p), & p_{i} \in (p_{i}, p_{i+1}].
\end{align*}
That is, in the region $[p_{i}, p_{i+1})$, $u[i]$ is a lower bound on $\overline{F}^{-1}$ and $d[i]$ is an upper bound on $\underline{F}^{-1}$ in $(p_{i}, p_{i+1}]$. In the case that the p-box characterises a precise distribution and $\overline{F}^{-1}(p) = \underline{F}^{-1}(p)$ gives:  $u[i+1] = d[i]$. Figure \ref{fig:figure2} shows examples of a distribution and a p-box with various outer representations.

@@

  struct pbox{T <: Real}
    u :: Vector{T}    # left discrete inverse
    d :: Vector{T}    # right discrete inverse
    m :: Interval{T}  # interval mean
    v :: Interval{T}  # interval variance
    shape :: String   # shape
  end
\end{lstlisting}
\noindent where the vectors $u$ and $d$ are the "up" and "down", or the "left" and "right", edges of the p-box. There can be confusion since $\overline{F}$ is the upper bound on the cdf, but $\overline{F}^{-1}$ is the lower bound on the random variable, i.e. the left edge of the p-box. We therefore call $\overline{F}$ the "up" or "left" edge, and $\underline{F}$ the "down" or the "right" edge. The intervals $m$ and $v$ are the interval moments, and the $shape$ is a string containing any distribution family information. An example of a normal distribution constructor is as follows
\begin{lstlisting}[language = Julia]
  using Distributions
  const steps = 200

@@

    return interval(lb, ub)
  end
\end{lstlisting}
\noindent Other available functions on p-boxes which we do not describe here are
\begin{itemize}
  \item \textbf{env} or $\cup$: takes the envelope or hull of two p-boxes. A convex set-union of the probability sets,
  \item \textbf{imp} or $\cap$: imposition or intersection of two p-boxes. Set-intersection of probability sets,
  \item \textbf{subseteq} or $\subseteq$: checks if a p-box is a subset or equal to another,
  \item \textbf{in} or $\in$: checks if a distribution is a member of a p-box,
  \item \textbf{mixture}: a stochastic mixture of intervals or p-boxes.
\end{itemize}
\subsection{Where do p-boxes come from?}
\noindent In this section we discuss some situations where p-boxes arise naturally.
\subsubsection{Partial distributional information}
In situations where the parametric family of the uncertainty is known, for example it is known that $X$ is normally distributed $X \sim N(\mu_{X}, \sigma_{X}^{2})$ or that $Y$ is a uniform $Y \sim U(a_{Y},b_{Y})$, but the precise parameters to these functions are unknown, then p-boxes are a natural representation. In the case that bounds are known on parameters, then bounds on the cdf can be found using the interval parameters and the inverse cdf of the parametric family. Figure \ref{fig:figure3} shows a variety of p-boxes constructed this way.
\subsubsection{Missing family but partial moment information}
\label{section:Moments}
It is well known that generally a random variables distribution function cannot be uniquely determined solely from its moment information, even with an infinite number of moments, unless some assumptions are made about the family (i.e. it's normal). In probability theory this is known as the \textit{moment problem}. However, in these situations it is possible to define a p-box which bounds all random variables with the known properties, without loss of generality or making unwarranted assumptions.

These p-boxes are constructed using the Chebyshev (mean and variance)~\cite{chebyshev1874valeurs}, Markov (mean and one bound)~\cite{markoff1900question} and Cantelli (mean and range) inequalities from probability theory. These inequalities allow for the cdf to be bounded from various range and moment information. For example, the Chebyshev inequality can be used to find p-box bounds from the random variables mean and variance
\begin{equation*}
  \underline{F}_{X}(x) = \begin{cases} 1 / (1 + (x - \mu_{X})^2 / \sigma^{2}_{X}),& \text{if } \mu_{X}<x \\
    1,& \text{if } x\leq \mu_{X} \end{cases}
\end{equation*}
\begin{equation*}
  \overline{F}_{X}(x)  = \begin{cases} 1 / (1 + \sigma^{2}_{X}/(x - \mu_{X})^2),& \text{if } x<\mu_{X} \\
  1,& \text{if } \mu_{X}\leq x \end{cases}
\end{equation*}
\noindent A list of available distribution-free constructors are
\begin{itemize}
  \item \textbf{MeanVar (or MeanStd)}: using Chebyshev,
  \item \textbf{MeanMin (or MeanMax)}: using Markov,
  \item \textbf{MeanMinMax}: using Cantelli,
  \item \textbf{MinMaxMeanVar (or MinMaxMeanStd)}: using~\cite{ferson2021distribution}
\end{itemize}
\noindent All distribution-free constructors may also take interval moments parameters.
\subsubsection{Operations involving intervals and distributions}
An arithmetic operation involving an interval and a precise distribution will produce set of distributions, i.e. a unique distribution for every real number in the interval, whose p-box bounds can be found. The following function performs a binary operation between a p-box and an interval
\begin{lstlisting}[language = Julia]
  using MomentArithmetic

@@

A method which is growing in popularity are inferential models~\cite{martin2015inferential}. These generalise classical inference methods to account for limited statistical knowledge, and give a statistical inference theory which is non-additive (as opposed to classical probability which is additive). These inference procedures, often based on random sets, give certain validity and performance guarantees. Inferential models can viewed from the perspective of bounding sets of distributions, and is closely linked to possibility theory~\cite{liu2020inferential}. A p-box can be prescribed to such structures, leading to structures like confidence boxes or c-boxes~\cite{fersoncomputing}, which follow the same arithmetic rules as p-boxes.

@@
